Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-10-06 Phase_II/draft/response.txt

|  Reviewer:
|  Dear Editor,
|  Thank you for sending me this manuscript by Rai et al on image
|  reconstruction from intensity interferometry. It addresses an important
|  topic, and may after major revision be suitable for publication in AJ. At a
|  high level, the paper does not demonstrate that the GAN presented actually
|  recovers much of the information about the stars presented; or that it
|  would be able to reconstruct images of other objects or with different uv
|  sampling.
|  Kind regards,
|  The Reviewer
|  
|  Comments
|
|  - Rapidly rotating stars can be fit analytically to these data. It may be
|  worth considering an example of an astrophysical source that would be more
|  interesting to fit, where the data analysis problem is less analytically
|  tractable, such as for complex scenes (spotted stars, or disks, for
|  instance).

Counter-argument: We can justify the choice of gravity-darkened stars.

1. Other image-reconstruction methods have been used with Michelson
   interferometry for this type of source.

The response to this in the Introduction:

Image reconstruction in gravity darkened fast rotating stars has long been examined using various methods in MI \citep{vanBelle2001, DomicianodeSouza2003, DomicianodeSouza2005, Monnier2007, Pedretti2009, Zhao2009, Martinez2021}.

@@

2. SII has recently measured ellipticity, so gravity-darkening is a
   natural next step.

The response to this in the Introduction:

Recently photosphere oblateness of $\gamma$-Cassiopeia \citep{2025arXiv250615027A} has been measured at the VERITAS observatory using II. These results, and especially, that by Archer et al. \cite{2025arXiv250615027A}  put our work in context. Thus our simulation is a natural next step of the work of Archer et al and others.

@@

|  - On the second page, there is quite a long discussion of the problem of
|  phase retrieval, including many citations. This is good, though perhaps a
|  little long and could be condensed. But there needs to be a clearer
|  statement of why phase retrieval matters - the Fourier transform is
|  Hermitian, and the consequence is that image modes that have even symmetry
|  under inversion are encoded strictly in amplitudes, and with odd symmetry
|  under inversion in phases. The result is that it intensity interferometry
|  strictly contains no information about odd modes of the image and these
|  must be constrained by the even modes through regularization/priors. This
|  idea needs to be discussed.

Text has been revised.

|  - I think we are seeing the Hermitian symmetry issue in Figure 13: while
|  the GANs do reasonable reconstructions of the equator-on gravity-darkened
|  stars, in the third and separately in the final examples, you see a pole-on
|  star with significant asymmetry in the ground-truth image, which the GANs
|  completely miss. In this figure, the Difference image is illegible; you
|  should try to unravel this into a residuals plot in 1D by mapping those
|  pixelized residuals back to baselines, and where differences are used in
|  images, use a diverging colourmap like coolwarm or seismic rather than
|  viridis.

Asymmetry is recovered in some cases: investigate!
We increased the number of training, validation and testing dataset. 
Also we increase the training time from 60k to 100k and result has 
been improvised. The difference plot is now in 1-D.

### 2025-10-06 Phase_II/draft/ParaGan.tex
ParaGan.tex, main.tex, response.txt

The process to train the model is summarized here:
\begin{figure*}[htbp]
	\centering
	\begin{tikzpicture}[node distance=0.9cm]

		% Data pipeline
		\node[block, fill=blue!15] (data) {Ground Truth \\ (Create $N$ Images of Stellar Objects)};
		\node[block, fill=red!10, below=of data] (pre) {Simulation \\ (Intensity Interferometry (II) Data)};
		\node[block, fill=green!20, below=of pre, minimum width=5cm] (split) {Input Dataset (Merge of Ground Truth \\ and Simulation of II) Split \\ (Training, Validation, Testing)};

		% Split branches
		\node[block, fill=green!15, below left=0.6cm and 2.5cm of split] (train) {Training Set \\ (Learn Model Parameters)};
		\node[block, fill=yellow!20, below=0.6cm of split] (val) {Validation Set \\ (Tune Hyperparameters)};
		\node[block, fill=orange!20, below right=0.6cm and 2.5cm of split] (test) {Testing Set \\ (Final Evaluation)};

		% Model training big box
		\node[block, fill=orange!20, below=3cm of val, minimum width=10cm, minimum height=1.2cm] (mod) {Train the Model \\ (Generator + Discriminator)};

		% Generator side (left)
		\node[block, fill=cyan!15, below=1.2cm of mod, xshift=-3.5cm] (noise) {Random Noise $z$};
		\node[block, fill=green!20, below=of noise] (gen) {Generator $G(z)$};
		\node[block, fill=blue!10, below=of gen] (fake) {Generated Image};
		\node[block, fill=purple!20, below=of fake] (gloss) {Generator Loss \\ (Update $G$)};

		% Discriminator side (right)
		\node[block, fill=yellow!20, below=1.2cm of mod, xshift=+3.5cm] (real) {Real Image \\ (From Training Set)};
		\node[block, fill=orange!20, below=of real] (disc) {Discriminator $D(x)$};
		\node[block, fill=red!20, below=of disc] (out) {Output: Real/Fake};
		\node[block, fill=purple!30, below=of out] (dloss) {Discriminator Loss \\ (Update $D$)};

@@

		% Connections for data pipeline
		\draw[arrow] (data) -- (pre);
		\draw[arrow] (pre) -- (split);
		\draw[arrow] (split.south west) |- (train.north);
		\draw[arrow] (split.south) -- (val.north);
		\draw[arrow] (split.south east) |- (test.north);
		\draw[arrow] (train.south) |- (mod.north);

		% GAN arrows
		\draw[arrow] (mod) -- (noise);
		\draw[arrow] (mod) -- (real);
		\draw[arrow] (noise) -- (gen);
		\draw[arrow] (gen) -- (fake);
		\draw[arrow] (fake.east) -- ++(0.5,0) |- (disc.west);
		\draw[arrow] (real) -- (disc);
		\draw[arrow] (disc) -- (out);

		% Feedback loops
		\draw[arrow] (out.south west) -- ++(-3,0) |- (gloss.east);
		\draw[arrow] (gloss.west) -- ++(-3,0) |- (gen.west);
		\draw[arrow] (out.south east) -- ++(3,0) |- (dloss.east);
		\draw[arrow] (dloss.east) -- ++(3,0) |- (disc.east);

@@

	\end{tikzpicture}
	\caption{The cGAN-based image reconstruction block diagram. Data is split into training (used for GAN optimization), validation (for hyperparameter tuning), and testing (for final evaluation). The GAN training loop optimizes Generator and Discriminator, while validation and testing feed into performance evaluation metrics.}
	\label{fig:gan_block}
\end{figure*}

>>>>

The process to train the model is summarized here:
\iffalse
\begin{figure*}[htbp]
	%\centering
	\begin{center}
		\begin{tikzpicture}[node distance=0.7cm]

			% Data pipeline
			\node[block, fill=blue!15] (data) {Ground Truth \\ (Create SS Images of Stellar Objects)};
			\node[block, fill=red!10, below=of data] (pre) {Simulation \\ (Intensity Interferometry (II) Data)};
			\node[block, fill=green!20, below=of pre, minimum width=5cm] (split) {Input Dataset (Merge of Ground Truth \\ and Simulation of II) Split \\ (Training, Validation, Testing)};

			% Training, Validation, Testing
			\node[block, fill=green!15, below left=0.6cm and 1.0cm of split] (train) {Training Set \\ (Learn Model Parameters)};
			\node[block, fill=green!10, below=of split] (val) {Validation Set \\ (Tune Hyperparameters)};
			\node[block, fill=green!05, below right=0.6cm and 1.0cm of split] (test) {Testing Set \\ (Final Evaluation)};

			% Model training block
			\node[block, fill=orange!20, below= 1.0cm of val, minimum width=10cm, minimum height=1.2cm] (mod) {Train the Model \\ (Generator + Discriminator)};

			% Generator side (left)
			\node[block, fill=yellow!15, below=2.2cm of mod, xshift=-3.5cm] (noise) {Random Noise $z$};
			\node[block, fill=blue!10, below=of noise] (gen) {Generated Image $\hat{x}$};
			\node[block, fill=purple!20, below=of gen] (gloss) {Generator Loss \\ (Update $G$)};

			% Discriminator side (right)
			\node[block, fill=yellow!20, below=2.2cm of mod, xshift=+3.5cm] (real) {Real Image \\ (From Training Set)};
			\node[block, fill=orange!20, below=of real] (disc) {Discriminator $D(x)$};
			\node[block, fill=red!20, below=of disc] (out) {Discriminator Output};
			\node[block, fill=purple!30, below=of out] (dloss) {Discriminator Loss \\ (Update $D$)};

			% Final result block
			\node[block, fill=teal!20, below= 3cm of val, minimum width=5cm] (res) {Result / Evaluation \\ (Performance Metrics: Accuracy)};

			% Connections
			\draw[arrow] (data) -- (pre);
			\draw[arrow] (pre) -- (split);
			%\draw[arrow] (split.south west) |- (train.north);
			\draw[arrow] (split.west) -| (train.north);
			\draw[arrow] (split.south) -- (val.north);
			\draw[arrow] (split.east) -| (test.north);
			\draw[arrow] (train.south) -- (mod.north);

			%Newly Introduced Connections (S Sarangi)
			\draw[arrow] (val.south) -- (mod.north);
			\draw[arrow] (mod.south) -- (res.north);

			% GAN arrows
			\draw[arrow] (mod.south) -- (noise.north);
			\draw[arrow] (mod.south) -- (real.north);
			\draw[arrow] (noise.south) -- (gen.north);
			\draw[arrow] (gen.south) -- (gloss.north);
			\draw[arrow] (real.south) -- (disc.north);
			\draw[arrow] (gen.east) -- ++(0.5,0) |- (disc.west);
			\draw[arrow] (disc.south) -- (out.north);
			\draw[arrow] (out.south) -- (dloss.north);

			% Validation & Testing to Result
			%\draw[arrow] (val.south) -- (res.north);
			\draw[arrow] (test.south) -- (res.north);

		\end{tikzpicture}
	\end{center}
\caption{A block diagram for the process of image reconstruction through GAN.}
\end{figure*}
\fi

\iffalse
\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{arrows, arrows.meta, positioning}

\tikzset{
  block/.style = {draw, rectangle, minimum height=1.5em, minimum width=3em, align=center},
  arrow/.style = {->, thick},
}

%\begin{document}
\fi

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[node distance=0.9cm]
%\begin{tikzpicure}[scale = 0.5]

% Data pipeline
\node[block, fill=blue!15] (data) {Ground Truth \\ (Create SS Images of Stellar Objects)};
\node[block, fill=red!10, below=of data] (pre) {Simulation \\ (Intensity Interferometry (II) Data)};
\node[block, fill=green!20, below=of pre, minimum width=5cm] (split) {Input Dataset (Merge of Ground Truth \\ and Simulation of II) Split \\ (Training, Validation, Testing)};

% Training, Validation, Testing
\node[block, fill=green!15, below left=0.6cm and 2.5cm of split] (train) {Training Set \\ (Learn Model Parameters)};
\node[block, fill=green!10, below=of split] (val) {Validation Set \\ (Tune Hyperparameters)};
\node[block, fill=green!05, below right=0.6cm and 2.5cm of split] (test) {Testing Set \\ (Final Evaluation)};

% Model training block
\node[block, fill=orange!20, below=3cm of val, minimum width=10cm, minimum height=1.2cm] (mod) {Train the Model \\ (Generator + Discriminator)};

% Generator side (left)
\node[block, fill=yellow!15, below=1.2cm of mod, xshift=-3.5cm] (noise) {Random Noise $z$};
\node[block, fill=blue!10, below=of noise] (gen) {Generated Image $\hat{x}$};
\node[block, fill=purple!20, below=of gen] (gloss) {Generator Loss \\ (Update $G$)};

% Discriminator side (right)
\node[block, fill=yellow!20, below=1.2cm of mod, xshift=+3.5cm] (real) {Real Image \\ (From Training Set)};
\node[block, fill=orange!20, below=of real] (disc) {Discriminator $D(x)$};
\node[block, fill=red!20, below=of disc] (out) {Discriminator Output};
\node[block, fill=purple!30, below=of out] (dloss) {Discriminator Loss \\ (Update $D$)};

@@

% Connections
\draw[arrow] (data) -- (pre);
\draw[arrow] (pre) -- (split);
\draw[arrow] (split.south west) |- (train.north);
\draw[arrow] (split.south) -- (val.north);
\draw[arrow] (split.south east) |- (test.north);
\draw[arrow] (train.south) -- (mod.north);

% GAN arrows
\draw[arrow] (mod.south) -- (noise.north);
\draw[arrow] (mod.south) -- (real.north);
\draw[arrow] (noise.south) -- (gen.north);
\draw[arrow] (gen.south) -- (gloss.north);
\draw[arrow] (real.south) -- (disc.north);
\draw[arrow] (gen.east) -- ++(0.5,0) |- (disc.west);
\draw[arrow] (disc.south) -- (out.north);
\draw[arrow] (out.south) -- (dloss.north);

@@

\end{tikzpicture}
\end{figure}

%\end{document}
